---
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
fontsize: 11pt
header-includes:
- \usepackage{placeins}
- \usepackage{fancyhdr}
- \usepackage{pdflscape}
- \usepackage{gensymb}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}

bibliography: C:/Users/mypath/myproject/data/references.bib
#bibliography: .data/references.bib                   # Use relative filepaths whenever possible! OS is friendlier than PC
#bibliography: [./data/ch1.bib,./data/ch2.bib]        # on OS, you can use multiple bibliographies like so

csl: C:/Users/mypath/myproject/data/the-auk.csl
#csl stylizes your .bib file to match a journal of your choice
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 options(scipen=1, digits=2)                                      #this controls digits when rounding
```

\pagenumbering{gobble}

```{r child = '02_Titlepage.Rmd'}
```

\newpage
\pagestyle{fancy}

\fancyhead[LE,RO]{}
\fancyhead[LO,RE]{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

\pagenumbering{roman}

\fancyhead[CO,CE]{Table of Contents}
\setcounter{tocdepth}{2}
\tableofcontents


\newpage
\fancyhead[CO,CE]{Dissertation Abstract}
\addcontentsline{toc}{section}{Dissertation Abstract}
\section*{Abstract}
```{r child = '03_Abstract.Rmd'}
```

\newpage
\pagenumbering{arabic}
\fancyhead[CO,CE]{Chapter 1}

```{r child = '04_Chapter1.Rmd'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Chapter 2}

```{r child = '05_Chapter2.md'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Literature Cited}
\section*{Literature Cited}

```{r child = '06_LitCited.R'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Supplemental Information}

```{r child = '07_Supplemental.Rmd'}
```


