---
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
fontsize: 11pt
header-includes:
- \usepackage{placeins}
- \usepackage{fancyhdr}
- \usepackage{pdflscape}
- \usepackage{gensymb}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
bibliography: [../Chapter 1/ch1_references.bib,ch2.bib,ch3.bib] #C:/Users/Julie/Box Sync/R/Chapter 1/ch1_references.bib
csl: the-auk.csl #C:/Users/Julie/Box Sync/R/Chapter 1/the-auk.csl
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 options(scipen=1, digits=2)
```

\pagenumbering{gobble}

```{r child = 'titlepage.Rmd'}
```

\newpage
\pagestyle{fancy}

\fancyhead[LE,RO]{}
\fancyhead[LO,RE]{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

\pagenumbering{roman}

\fancyhead[CO,CE]{Table of Contents}
\setcounter{tocdepth}{2}
\tableofcontents



```{r child = 'abstract.Rmd',eval=FALSE}
#move the contents outside the chunk when ready to include
\newpage
\fancyhead[CO,CE]{Abstract}
\addcontentsline{toc}{section}{Abstract}
\section*{Abstract}
```

\newpage
\pagenumbering{arabic}
\fancyhead[CO,CE]{Chapter 1}

```{r child = '../Chapter 1/chapter1.Rmd'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Chapter 2}

```{r child = 'chapter2.Rmd'}
```


\FloatBarrier
\newpage
\fancyhead[CO,CE]{Chapter 3}

```{r child = 'chapter3.md'}
```

\FloatBarrier
\newpage
\fancyhead[CO,CE]{Literature Cited}
\section*{Literature Cited}

```{r child = 'lit_cited.R'}
#<div id="refs"></div>
```


\FloatBarrier
\newpage
\fancyhead[CO,CE]{Supplemental Information}

```{r child = 'supplemental.Rmd'}
```
